Homogeneous Light Curves for Stars in Clusters from TESS: Method \& Sector 1-4 Data Release

### 2019-02-01 paper/ms.tex
minor paper improvements

\title{Homogeneous Light Curves for Stars in Clusters from TESS}

@@

We reduced the TESS images to lightcurves by performing a sequence of
steps using stand-alone programs.  A conceptual overview of our
``pipeline'' is given in Figure~\ref{fig:pipeline}.  Similar pipelines
have been presented by \citet{Pal_2009},
\citet{soares-furtado_image_2017} and \citet{oelkers_precision_2018}.

We begin with the calibrated full frame images produced by SPOC
(\S~\ref{subsec:observations}).  We then perform a collection of
preparatory steps, including source extraction of bright stars,
astrometry using the resulting positions, and coarse simple aperture
photometry (\S~\ref{subsec:preparation}).  Using the shape values from
the initial astrometry, we select an astrometric reference frame and
transform all of the calibrated images to it.  We construct a
photometric reference by stacking a collection of frames, and then
convolve all the transformed frames to match the photometric reference,
and subtract (\S~\ref{subsec:imagesubtraction}).  We perform aperture
photometry on the subtracted images using positions projected onto the
frame from Gaia DR2.  We detrend the resulting lightcurves
(\S~\ref{subsec:lcdetrending}) in two subsequent steps.  First, we apply
External Parameter Decorrelation (EPD, \citealt{bakos_2010}), and then
we apply a Trend Filtering Algorithm (TFA, \citealt{kovacs_trend_2005}).
Finally, we assess the resulting lightcurves using a variety of
statistics for their white noise and red noise properties
(\S~\ref{sec:results}).

@@

For us, the main data product of interest is the calibrated full frame
image (FFI).  Each TESS camera reads out every 2 seconds.  To produce
a manageable telemetric load, the resulting pixel values are summed by
the onboard computer into 30 minute averages. (An on-board cosmic ray
mitigation algorithm is applied.) Once transmitted to the ground, the
raw images are calibrated by the Science Processing Operations Center.
The calibration process includes an overscan, bias, and dark current
correction, and also dividing out a flat field.  The details are given by
\citet{clarke_kepler_2017}, and the resulting science data products are
described by \citet{tess_data_product_description_2018}.

@@

Next, we perform some initial analysis steps to produce metadata
needed during image subtraction.  To obtain an astrometric solution
(independent from the WCS data packaged with the frames), we use
\texttt{fistar} to perform source extraction on bright stars in each
image.  We pass the resulting source lists through
\texttt{astrometry.net} \citep{lang_2010}, which returns an
astrometric solution for each frame packaged in the WCS format
\citep[][Sec.~8]{pence_fits_2010}.  During the initial source
extraction, we also fit elongated gaussians to the bright stars,
yielding the shape parameters (S,D,K), where the flux as a function of
position is assumed to take the form
\begin{align}
  f_{\rm elong}(\vec{x}) &= B + A \exp \{ -0.5 \times ( 
    S(\Delta x^2 + \Delta y^2) + \\
    &D(\Delta x^2 - \Delta y^2) +
    K(2\Delta x \Delta y)
  )  \},
\end{align}
for $\Delta x = x-x_0$, and $\Delta y = y - y_0$.  $S$ is typically
called the ``sharpness''.  For a nearly circular shape profile, ${\rm
FWHM} \approx 2.35\sqrt{S}$ \citep[{\it e.g.},][]{Pal_2009}.  These
shape parameters are later used when selecting an astrometric
reference (\S~\ref{subsec:imagesubtraction}).
We note that the fast focal ratio of the TESS cameras introduces
significant comatic aberrations to the images, stars closer to the
center of the field are more round; stars towards the edges are more
triangular.

With the resulting WCS information, we then project a source catalog
from Gaia-DR2 onto the frame, down to a pre-selected magnitude cutoff
\citep{gaia_collaboration_gaia_2018}.  We use these expected positions
to center the apertures in our photometry, rather than attempting to
detect the positions.  Such ``forced-aperture photometry'' is
preferable to performing source extraction because of the large TESS
pixels, and the accurate Gaia positions.  The Gaia-DR2 epoch is
J2015.5, so even the fastest-moving stars with proper motions of
$\sim$$1\,{\rm arcsecond}\,{\rm yr}^{-1}$ are still well within one pixel
of their predicted positions in the TESS images.  The projection and
catalog-indexing is performed using
\texttt{gaia2read}\footnote{\url{github.com/samuelyeewl/gaia2read}}
(CITE Jason Kim junior thesis).

@@

@@

\begin{figure*}[t]
    \begin{center}
		\includegraphics[height=0.97\textheight]{f4.png}
    \end{center}
    \vspace{-0.8cm}
    \caption{
      {\bf Flux as a function of ``external parameters'' for four
      representative stars.} The left two columns are stars at the
      corner of a camera's field; the right two columns are from the
      centers.
      Each row shows a different parameter along the $x$-axis, given
      in text at the bottom right of each subplot.
      ``Hooks'' are common features in flux as a function of
      temperature and centroid position.
      \S~\ref{sec:flux_vs_external_parameters} gives a verbose
      description.
     \label{fig:flux_vs_external_parameters}
    }
\end{figure*}

The traditional approach to EPD 
\citep[][]{Pal_2009,bakos_2010,zhang_precision_2016} is to fit and
subtract a model of the form

@@

where $\vec{\theta}$ is a vector of parameters such as the shape
parameters $(s,d,k)$, their products $(s^2, s\cdot d, d^2, \ldots)$,
the temperature $T$ of the CCDs\footnote{ We used the temperature from
the on-chip aluminum-copper sensor measurements included in the
engineering
data\footnote{\url{archive.stsci.edu/missions/tess/engineering/}}.  },
the centroid positions $(x,y)$, the fractional part of the
centroid positions $(\{x\},\{y\})$, and any other parameters that are
expected\footnote{ The fractional centroid positions might matter
because intra-pixel quantum efficiency variations could affect the
measured stellar brightness.  The varying temperature $T$ of the CCD
electronics might matter.  } to correlate strongly with the observed
flux.  The coefficients $c_i$ are calculated through linear
least-squares, and subtracted to produce ``EPD'' lightcurves.

The premise of this model is that the correlations between the
magnitudes and these external parameters are linear.  For ground-based
CCD data (e.g., HATNet, HATS, and Nikon DSLRs), \citet{bakos_2010} and
\citet{zhang_precision_2016} have verified that this model is a
reasonable fit to the data.

To discern whether such a model describes the TESS data, we examined scatter 
plots of each
parameter, as a function of all the other parameters.  We also
examined plots of each parameter as a function of time.
A few prominent trends were present.

@@

Given the characteristics of the variability, a linear model of the
form given in Equation~\ref{eq:classical_EPD} is not applicable.  To
fit out the correlations between flux and parameters which most
commonly exhibited ``hooks'', we explored fitting a parametric open
curve (an $N$-dimensional B-spline, \citealt{dierckx_curve_1996}) to
the flux, centroid positions, and temperatures simultaneously.  We
selected the number of knots through brute-force, by calculating
$\chi^2$ for the model fit of a grid of knots, and minimizing the
Bayesian Information Criterion.  Though this approach showed some
initial promise, even with ``optimal'' knot-selection (in the BIC
sense) it introduced undesirable residuals in the lightcurves, and
also distorted transits.

>>>>

\title{Homogeneous Light Curves for Stars in Clusters from TESS:
Method \& Sector 1-4 Data Release}

@@

We reduced the TESS images to lightcurves by performing a sequence of
steps using stand-alone programs.  A conceptual overview of our
``pipeline'' is given in Figure~\ref{fig:pipeline}.  
Our overall method is in the spirit of the reduction approaches used
by \citet{Pal_2009}, \citet{soares-furtado_image_2017} and
\citet{oelkers_precision_2018}.

We begin with the calibrated full frame images produced by SPOC
(\S~\ref{subsec:observations}).  We then perform a collection of
preparatory steps, including source extraction of bright stars,
astrometry using the resulting positions, and coarse simple aperture
photometry (\S~\ref{subsec:preparation}).  Using the shape values from
the initial astrometry, we select an astrometric reference frame and
transform all of the calibrated images to it.  We construct a
photometric reference by stacking a collection of frames, and then
convolve all the transformed frames to the photometric reference,
and subtract (\S~\ref{subsec:imagesubtraction}).  We perform aperture
photometry on the subtracted images using positions projected onto the
frame from Gaia DR2.  We detrend the resulting lightcurves
(\S~\ref{subsec:lcdetrending}), and then assess the resulting
lightcurves using a variety of statistics for their white noise and
red noise properties (\S~\ref{sec:results}).

@@

For us, the main data product of interest is the calibrated full frame
image (FFI).  Each TESS camera reads out every 2 seconds.  To produce
a manageable telemetric load, the resulting pixel values are averaged
by the onboard computer into 30 minute exposures. (An on-board cosmic
ray mitigation algorithm is applied.) Once transmitted to the ground,
the raw images are calibrated by the Science Processing Operations
Center.  The calibration process includes an overscan, bias, and dark
current correction, and also dividing out a flat field.  The details
are given by \citet{clarke_kepler_2017}, and the resulting science
data products are described by
\citet{tess_data_product_description_2018}.

@@

Next, we perform some initial analysis steps to produce metadata
needed during image subtraction.  To obtain an astrometric solution
(independent from the WCS data packaged with the frames), we use
\texttt{fistar} to perform source extraction on bright stars in each
image.  We pass the resulting source lists through
\texttt{astrometry.net} \citep{lang_2010}, which returns an
astrometric solution for each frame packaged in the WCS format
\citep[][Sec.~8]{pence_fits_2010}.  During the initial source
extraction, we also fit elongated gaussians to the bright stars,
yielding the shape parameters $(s,d,k)$, where the flux as a function
of position is assumed to take the form
\begin{align}
  f_{\rm elong}(\vec{x}) &= B + A \exp \{ -0.5 \times ( 
    s(\Delta x^2 + \Delta y^2) + \\
    \nonumber
    &d(\Delta x^2 - \Delta y^2) +
    k(2\Delta x \Delta y)
  )  \},
\end{align}
for $\Delta x = x-x_0$, and $\Delta y = y - y_0$.  For a nearly
circular shape profile, the sharpness $s$ is related to the FWHM as
${\rm FWHM} \approx 2.35\sqrt{s}$ \citep[{\it e.g.},][]{Pal_2009}.
These shape parameters are later used when selecting an astrometric
reference (\S~\ref{subsec:imagesubtraction}).  We note that the fast
focal ratio of the TESS cameras introduces significant comatic
aberrations to the images: stars closer to the center of the field are
more round, while stars towards the edges are more triangular.

With the resulting WCS information, we then project a source catalog
from Gaia-DR2 onto the frame, down to a pre-selected magnitude cutoff
\citep{gaia_collaboration_gaia_2018}.  We use these expected positions
to center the apertures in our photometry, rather than attempting to
detect the positions.  Such ``forced-aperture photometry'' is
preferable to performing source extraction because of the large TESS
pixels, and the accurate Gaia positions.  The Gaia-DR2 epoch is
J2015.5, so even the fastest-moving stars with proper motions of
$\sim$$1\,{\rm arcsecond}\,{\rm yr}^{-1}$ are still well within one pixel
of their predicted positions in the TESS images.  The projection and
catalog-indexing is performed using
\texttt{gaia2read}\footnote{\url{github.com/samuelyeewl/gaia2read}}
\citep{kim_2018_gaia2read}.

@@

The core operation of ``classical'' image subtraction
attempts to match a photometric reference $R$ and a target image $I$ by 
computing a convolution kernel.
The kernel, once applied to the high signal-to noise reference,
produces a model image, $M_{xy}$,
\begin{equation}
    M_{xy} = (R \otimes K)_{xy} + B_{xy},
\end{equation}
where $B_{xy}$ is a component of the model image that allows for
background variations.
The convolution kernel $K$ is typically decomposed onto a
basis, $K = \sum_i c_i K_i$, and the coefficients are found
by minimizing
\begin{equation}
    \chi^2 = \sum_{xy} \left( \frac{I_{xy} - M_{xy}}{\sigma_{xy}} \right)^2,
\end{equation}
where $\sigma_{xy}$ is the uncertainty in the target image pixel values.
Photometry is then performed on the difference image $D_{xy}$, where
$D_{xy} = I_{xy} - M_{xy}$.

% FIXME: describe gaussian kernel (used in e.g., KELT, and maybe other 
%pipelines?). Perhaps give Eqns for delta kernel.
The choice of kernel ...

@@

\begin{figure*}[t]
    \begin{center}
		\includegraphics[height=0.95\textheight]{f4.png}
    \end{center}
    \vspace{-0.8cm}
    \caption{
      {\bf Flux as a function of ``external parameters'' for four
      representative stars.} The left two columns are stars at the
      corner of a camera's field; the right two columns are from the
      centers.
      Each row shows a different parameter along the $x$-axis, given
      in text at the bottom right of each subplot.
      ``Hooks'' are common features in flux as a function of
      temperature and centroid position.
      \S~\ref{sec:flux_vs_external_parameters} gives a verbose
      description.
     \label{fig:flux_vs_external_parameters}
    }
\end{figure*}

The traditional approach to EPD 
\citep[][]{Pal_2009,bakos_2010,zhang_precision_2016} is to fit and
subtract a model for the magnitudes $m$ of the form

@@

where $\vec{\theta}$ is a vector of parameters such as the shape
parameters $(s,d,k)$, their products $(s^2, s\cdot d, d^2, \ldots)$,
the temperature $T$ of the instrument or environment\footnote{ We used
the temperature from the on-chip aluminum-copper sensor measurements
included in the engineering
data\footnote{\url{archive.stsci.edu/missions/tess/engineering/}}.  },
the centroid positions $(x,y)$, the fractional part of the
centroid positions $(\{x\},\{y\})$, and any other parameters that are
expected\footnote{ The fractional centroid positions might matter
because intra-pixel quantum efficiency variations could affect the
measured stellar brightness.  The varying temperature $T$ of the CCD
electronics might matter.  } to correlate strongly with the observed
flux.  The coefficients $c_i$ are calculated through linear
least-squares, and subtracted to produce ``EPD'' lightcurves.

The premise of this model is that the correlations between the
magnitudes and the external parameters are linear.  For ground-based
CCD data (e.g., HATNet, HATS, and Nikon DSLRs), \citet{bakos_2010} and
\citet{zhang_precision_2016} have verified that this model is a
reasonable description to the data.

To discern whether such a model extends to the TESS data, we examined
scatter plots of each parameter, as a function of all the other
parameters.  We also examined plots of each parameter as a function of
time.  A few prominent trends were present.

@@

Given the characteristics of the variability, a linear model of the
form given in Equation~\ref{eq:classical_EPD} is not applicable.  To
fit out the correlations between flux and parameters which most
commonly exhibited ``hooks'', we explored fitting a parametric open
curve (an $N$-dimensional B-spline, \citealt{dierckx_curve_1996}) to
the flux, centroid positions, and temperatures simultaneously.  We
selected the number of knots through brute-force, by calculating
$\chi^2$ for the model fit over a grid of possible knots, and
minimizing the Bayesian Information Criterion.  Though this approach
showed some initial promise, even with ``optimal'' knot-selection (in
the BIC sense) it introduced undesirable residuals in the lightcurves,
and also distorted transits.
